---
title: "Tools for Reproducible Research in Linguistics"
subtitle: "LSA 2019"
author: "Bradley McDonnell^1^, Na-Rae Han^2^, Eve Koller^1^ <br/> <br/> ^1^University of Hawai‘i at Mānoa<br/> ^2^University of Pittsburg"
date: "2019/01/03 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r, load_refs, echo=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear", 
           cite.style = "authoryear", 
           style = "markdown",
           hyperlink = FALSE, 
           dashed = FALSE)
rrBib <- ReadBib("intro_workshop_files/rr.bib", check = FALSE)
```


---
# Introduction & Introductions

 

---
# Who are we?

1. Brad McDonnell
  
2. Na-Rae Han

3. Eve Koller

---
# Why are we offering this workshop? 

**Purpose:** Increase literacy 

---
# Who are you?

## Share the following:

- Name
- Research interests
- Reasons for taking the workshop


---

# What is Reproducible Research?

[...] *reproducibility* [...] in research means that the data on which publications are based are made available so that other scientists could ostensibly verify the results for themselves.

`r Citep(rrBib, "gawneforthcomingreflections")`

---

class: inverse, middle, center

# What are some reasons for practicing reproducible research?


---

# Tools

Reproducible research is increasingly possible with the development of new tools.

1. `git` allows us to keep track of the changes we make to code, data, etc.
2. GitHub allows us 
  - to keep track of these changes while collaborating with others,
  - share research with others
3. RStudio and Jyputer notebooks allows us ...  




???

The means to make research reproducible 

 

---

# Goals of the workshop


---

# Overview of the workshop

| Time | Topic | Presenters |
|------|-------|------------|
| 9:00 - 9:15 | Introductions | |
| 9:15 - 9:30 | Overview of workshop | Brad | 
| 9:30 - 10:45 | [Introduction to git](intro_to_git.md) | Na-Rae |
| **10:45 - 11:00** | **Break** | |
| 11:00 - 12:30 | [Linking git and GitHub](linking_git_and_github.md) | Na-Rae, Brad |
| **12:30 - 1:30** |  **Lunch** | |
| 1:30 - 1:45 | RStudio with R demo  | Brad |
| 1:45 - 2:00 | Jupyter Notebooks with Python demo  | Na-Rae |
| 2:00 - 3:15 | Break out groups with RStudio and Jupyter Notebooks | |
| 3:15 - 3:30 | Concluding remarks |Brad|


---

# Summary

.pull-left[
## What we're teaching...

  1. git
  1. GitHub
  1. Jupyter Notebooks
  1. RStudio
]

.pull-right[
## What we're ***not*** teaching...
  1. R
  1. Python
]
---

# References


```{r, 'refs', results='asis', echo=FALSE}
PrintBibliography(rrBib, start = 1, end = 2)
```


--- 

class: inverse, middle, center

# Questions?

---



