---
title: "The Typology of Causative-Applicative Syncretism"
author: "Bradley McDonnell"
date: "1/1/2019"
output: 
  ioslides_presentation:
    widescreen: true
bibliography: bib/causative-applicative.bib
---

<!--TASK: 

1. Knit slides to see what they look like. Then change...

output: ioslides_presentation

to...

output: 
  ioslides_presentation:
    widescreen: true
    smaller: true

Notice a difference? (I think it looks much better.)

2. add the bibliography "caus-appl.bib" for these slides. 

-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-data, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(leaflet) # for creating maps
library(wesanderson) # for custom colors
library(knitr) # for tables
library(kableExtra) # for tables

## Load datasets
wals_200_caus_appl <- read_tsv("data/wals_200_caus_appl.csv")
wals_200_syncretism <- read_tsv("data/wals_200_syncretism.csv")

## Define colors for maps
pal <- colorFactor(palette = wes_palette(name = "Zissou1")[c(2,5)],
                   domain = levels(wals_200_caus_appl$causative_111a)
                   )
pal2 <- colorFactor(palette = wes_palette(name = "Zissou1")[c(5)],
                    domain = levels(wals_200_caus_appl$reported)
                    )
```


## Goals

>1. An overview of causatives and applicatives in the world's languages.
>1. Worldwide distribution of causative-applicative syncretism.
>    - Distribution within a subset of the WALS 200 sample.
>1. Challenges & prospects creating a typology of causative-applicative syncretism.

<!-- Task 
Try to add > before each, item in the list above... what happens?  
-->

## Causatives{.flexbox .vcenter}


## WALS map for causatives [@song2013nonperiphrastic]

```{r eval=require('leaflet'), echo=FALSE}
causative_map <- wals_200_caus_appl %>%
  na.omit() %>%
  leaflet() %>%
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(causative_111a), 
                   weight = 1, 
                   fillColor = ~pal2(causative_111a),
                   fillOpacity = 0.75,
                   opacity = 1,
                   radius =  3, 
                   label=~as.character(name)
  ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~causative_111a,
            title = "Causatives",
            opacity = 1
  )
causative_map
```

## Causatives 

```{r caus-table}
wals_200_caus_appl %>%
  na.omit() %>%
  group_by(causative_111a) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html', 
        align = rep('c', 3),
        col.names = c("Causatives",
                      "Total (n)",
                      "Total (%)")
        )  %>%
kable_styling(bootstrap_options = "striped", font_size = 30) %>%
row_spec(0, font_size = 30)
```

## Applicatives


## WALS map for applicatives [@polinsky2013applicative]

```{r appl-map}
applicative_map <- wals_200_caus_appl %>%
  na.omit() %>%
  leaflet() %>% 
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(applicative_109a), 
                   weight = 1, 
                   fillColor = ~pal2(applicative_109a),
                   fillOpacity = 0.75,
                   opacity = 1,
                   radius = 3, 
                   label=~as.character(name)
  ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~applicative_109a,
            title = "Applicatives",
            opacity = 1
  )
applicative_map
```


## Applicatives totals

```{r, echo=FALSE}
wals_200_caus_appl %>%
  na.omit() %>%
  group_by(applicative_109a) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html',
        align = rep('c',3),
        col.names = c("Applicatives",
                      "Total (n)",
                      "Total (%)")
        )  %>%
  kable_styling(bootstrap_options = "striped", font_size = 30) %>%
  row_spec(0, font_size = 30)
```


## Causative & Applicatives 

```{r caus-appl}
caus_appl_table <- wals_200_caus_appl %>%
  na.omit() %>%
  group_by(causative_111a, applicative_109a) %>%
  summarize(n=n()) 

caus_appl_table %>%
  kable(format = 'html', 
        align = rep('c',3),
        col.names = c("Causatives", 
                      "Applicatives", 
                      "Total (n)")
        )  %>%
  kable_styling(bootstrap_options = "striped", font_size = 30) %>%
  row_spec(0, font_size = 30) %>%
  collapse_rows(columns = 1, valign = "middle")
```


## Causative & Applicatives

```{r caus-apple-df, include = FALSE}
caus_appl_table_df <- caus_appl_table %>%
  ungroup %>%
  as.data.frame()
```

- Most langauges have a **causative** but *no* **applicative** <br/> (n = `r caus_appl_table_df[3,3]`)
- Many languages have a **causative** *and* an **applicative** <br/> (n = `r caus_appl_table_df[4,3]`)
- A small number of langauges have *neither* a **causative** *nor* an **applicative** <br/> (n = `r caus_appl_table_df[1,3]`)
- Very *few* langauges have an **applicative** but *no* **causative** <br/> (n = `r caus_appl_table_df[2,3]`)


## Causative-applicative syncretism map

```{r sync-map}
syncretism_map <- wals_200_syncretism %>%
  leaflet() %>%
  fitBounds(lng1 = -170, lat1 = -50, lng2 = 170, lat2 = 70) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(~longitude, 
                   ~latitude, 
                   color = ~pal(syncretism), 
                   weight = 1, 
                   radius = 4, 
                   fillColor = ~pal(syncretism), 
                   fillOpacity = 0.75,
                   opacity = 1,
                   label=~as.character(name)
                   ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~syncretism,
            title = "Syncretism",
            opacity = 1
  )
syncretism_map
```

## Causative-applicative syncretism

```{r, echo=FALSE}
wals_200_syncretism %>%
  group_by(syncretism) %>%
  summarize(n=n()) %>%
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%")) %>%
  kable(format = 'html',
        align = rep('c',3),
        col.names = c("Syncretism", 
                      "Total (n)",
                      "Total (%)")
        )  %>%
  kable_styling(bootstrap_options = "striped", font_size = 30) %>%
  row_spec(0, font_size = 30)
```

---- 

<!-- Note that this how you create a slide without a title in case you're curious -->

Go to <https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html> for more ideas of what to do with your slides. 

For even more on **isoslides** go to <https://openscience.nuance-r.com/ioslides_presentation_format.html>

For `kable` tables, go to <https://haozhu233.github.io/kableExtra/awesome_table_in_html.html>

For another style of slides using R Markdown, see `xaringan` <https://github.com/yihui/xaringan>

If you are interested in bibliographies see <https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html>

If you want to number tables, see: 
<https://www.r-bloggers.com/r-markdown-how-to-number-and-reference-tables/>

<<<<<<< HEAD
If you wanted to look at potential article templates: 
<https://github.com/rstudio/rticles>

=======
>>>>>>> 63e4a44399bb48695ec9d1dbd19da4f6f5d955b5
## References
